Active Learning for SAT Solver Benchmarking

### 2022-10-08 paper/paper.tex
minor-fixes

\title{Active Learning for SAT solver Benchmarking}

@@

\begin{abstract}
  Experimentation, in particular the evaluation of the runtime of a concrete implementation, is one of the most important phases in the development of algorithms.
  This also applies to solutions of the SAT problem (propositional sat\-is\-fia\-bi\-li\-ty).
  While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
  Benchmark compilation selects representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
  In this paper, we present an active learning based dynamic benchmark selection approach that predicts the rank of a new solver among its competitors with minimum running time and maximum rank prediction accuracy.
  We evaluated this approach using the Anniversary track dataset from the 2022 SAT Competition. 
  Our dynamic instance selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}. 
  In this paper, we also discuss the importance of instance families in the selection process. 
  Our tool provides a reliable way for SAT solver engineers to efficiently determine the performance of a new SAT solver.

@@

\subsubsection{Traditional benchmark-set-selection.}
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

@@

\section{Active Learning for SAT solver Benchmarking}
\label{sec:main}
Our framework follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
%JB: as the algorithm is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rather than algorithm (and change the term consistently in text)
%TF: changed
After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
We \emph{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
%Note that the prediction target values may be also a subset of $\mathbb{R}$.
%Refer to the next sections for more details.

\begin{algorithm}
  \caption{Incremental benchmarking framework}
  \label{algALBenchmark}

  \KwIn{A pool of instances $\mathcal{I}$, a set of known solvers $\mathcal{A}$ with known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and a new solver $\hat{a} \notin \mathcal{A}$.}

>>>>

\title{Active Learning for SAT Solver Benchmarking}

@@

\begin{abstract}
  Benchmarking is one of the most important phases in the development of algorithms.
  This also applies to solutions of the SAT problem (propositional sat\-is\-fia\-bi\-li\-ty).
  While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
  Benchmark compilation selects representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
  In this paper, we present an active learning based dynamic benchmark selection approach that predicts the rank of a new solver among its competitors with minimum running time and maximum rank prediction accuracy.
  We evaluated this approach using the Anniversary track dataset from the 2022 SAT Competition. 
  Our dynamic instance selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}. 
  In this paper, we also discuss the importance of instance families in the selection process. 
  Our tool provides a reliable way for SAT solver engineers to efficiently determine the performance of a new SAT solver.

@@

\subsubsection{Traditional Benchmark Selection.}
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

@@

\section{Active Learning for SAT Solver Benchmarking}
\label{sec:main}
Our framework follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
%JB: as the algorithm is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rather than algorithm (and change the term consistently in text)
%TF: changed
After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
We \emph{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
%Note that the prediction target values may be also a subset of $\mathbb{R}$.
%Refer to the next sections for more details.

\begin{algorithm}
  \caption{Incremental Benchmarking Framework}
  \label{algALBenchmark}

  \KwIn{Instances $\mathcal{I}$, Solvers $\mathcal{A}$, Runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$}
  \KwIn{New Solver $\hat{a} \notin \mathcal{A}$}
